---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
title: MPs Parliament Attendance Data Preparation 
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
---


```{r setup, eval = TRUE}

library(dplyr)
library(magrittr)
library(stringr)
```

## 16th Loksabha Sessions

Session dates and attendance of all members taken from the official
[website](http://loksabha.nic.in).

```{r}
## return the number of weekdays between dates d1 and d2}
num_weekdays <- function(d1, d2){
    weekend  <- c("Saturday", "Sunday")
    all_days <- seq(d1, d2, "1 day") %>% weekdays
    sum(!(all_days %in% weekend))
}

num_weekdays_v <-Vectorize(num_weekdays)

```

I copied the session dates into a csv file. Two sessions (IV and XIV
sessions) had a break. We calculate the number of weekdays in each
session.


```{r}
session_dates <- here::here("data/16th", "loksabha_session_dates.csv")
session_dates <- readr::read_csv(session_dates)

session_dates$session %<>% as.roman %>% as.integer

session_dates$start %<>% lubridate::dmy(.)
session_dates$end   %<>% lubridate::dmy(.)

session_dates$session_length <- num_weekdays_v(session_dates$start,
                                               session_dates$end)

sessions_durations <- session_dates %>%
                      group_by(session) %>%
                      summarise_if(is.numeric, sum)




```

Adding up the durations of all the sessions, we know that Lok Sabha has worked for 310 days so far.
The attendance information is unfortunately made available in doc
format. There were 14 files (one for each session). I converted each doc
file to docx and used the docxtractr package to extract the tables. I
know I've to extract the 4th table from each document because I had
queried the table structure using the \(docx_describe_tbls\) from the
package.


```{r}

attendance  <- here::here("data/16th", paste0(1:14, ".docx")) %>%
               lapply(function(fname){
                                docxtractr::read_docx(fname) %>%
                                docxtractr::docx_extract_tbl(4)})

```


Stacking together the information from all sessions, I get a usable data
frame after having cleaned up the column names and fixed column classes.
Some of the column names are really clunky, so lets fix that.


```{r}
names(attendance) <- 1:14
all_sessions <- bind_rows(attendance, .id = "session")

all_sessions %<>% janitor::clean_names()

all_sessions %<>% dplyr::rename(mp_name = name_of_member,
                               days_attended =
                               no_of_days_member_signed_the_register)


```

We will use the name of the MP to join this data set to the data set 
containing attributes. so we will clean-up the names as well as other 
character columns. Let us read in the attributes data first so that we
can carry our the same clean-up operations on both files.


## Features

What we have is a bare bones data set that records the name of each MP,
their constituency and the number of days they attended the parliament
(or at least, signed the register). I will now add some features such as
gender, education etc. that I care about to this information. The source
for this information is PRS Legislative Research, specifically 
[this](http://www.prsindia.org/mptrack/16loksabha/). While the data
is well-formatted and this file also includes percent attendance, I
chose not to use it as it was not clear how this percent has been
calculated i.e., what was the denominator in each case. It was also
becoming a tad annoying to extract the dates hidden in the notes.


```{r}
mps_attributes <- here::here("data", "MPTrack-16.xls") %>%
                   readxl::read_excel(na = "N/A") %>% 
                   janitor::clean_names()

```


We can use skimr to take a quick look at the data to see if there is 
anything missing. The output of skimr doesn't play well with tex so 
I've skipped it here.

## Data clean-up

Fix the classes for some columns. For instance, the number of term and
the political party should be a factor. Also, get rid of some messy
columns we won't use.


```{r}
mps_attributes %<>% select(-notes)


```


The  tricky bit is this - we now have to join this set of
attributes with the data on attendance using name and/or 
constituency. There are so many ways in which this can go 
wrong - variation on spelling, using or not using titles etc. 
We will do the best we can! For starters, make everything 
lowercase, get rid of space and hope for the best. 
Also replace all commas with hyphens as they can create 
problems later on when we write the output.

First let us save the name so that we can tack it back later
on.

```{r}

attr_names <- mps_attributes$mp_name


titles <- c("dr.", "smt.", "prof.", "col.", "shri",
            ",", "@", ".", "(", ")", " ")
for(t in titles){
  mps_attributes %<>% mutate_if(is.character, 
                                str_remove_all, 
                                fixed(t, ignore_case = TRUE))
  all_sessions %<>% mutate_if(is.character, 
                              str_remove_all, 
                              fixed(t, ignore_case = TRUE))
}

mps_attributes %<>% mutate_if(is.character, str_to_lower)
all_sessions   %<>% mutate_if(is.character, str_to_lower)

mps_attributes %<>% mutate_if(is.character, str_trim)
all_sessions   %<>% mutate_if(is.character, str_trim)

mps_attributes$full_name <- attr_names


                              
                             
```

Also, manually fix up some spelling differences between the two files.

```{r}
## constituency as spelt in mp_attributes = as in all_sessions
mps_attributes$constituency %<>%
 factor %>%   
 forcats::fct_recode(ramanathapuram = "ramanthapuram",
                     mahesana = "mehesana",
                     kushinagar = "khushinagar",
                     krishnagiri = "krisnagiri",
                     bangaon = "bongaon")

levels(mps_attributes$constituency) %<>% c("nominatedanglo-indian")

## set this to nominatedanglo-indian
mps_attributes[mps_attributes$nature_of_membership == "nominated",
               "constituency"] <-  "nominatedanglo-indian"

```

Join time!


```{r}

attendance_with_attributes <- left_join(all_sessions,
                                        mps_attributes,
                                        by = "constituency")

```

Joining by constituency is not ideal because a constituency 
may have been represented by more than one MP during a term.
We want to join by the name of the MP but as these two data 
sets come from different sources, there are likely to be 
discrepancies between spellings, full/partial name, 
use of titles etc. So we will join by constituency, tease out 
the rows where MP names are different and *manually* check if
they represent the same name. It has to be done manually 
because there are infinite ways in which a name could be 
written differently.

```{r n_conflict}
names_conflict <- here::here("data/16th",
                             "conflicting_names.csv")
attendance_with_attributes %>%
    filter(mp_name.x != mp_name.y) %>%
    select(mp_name.x, mp_name.y) %>%
    filter(!duplicated(mp_name.x)) %>%
    readr::write_csv(path = names_conflict)

```

There were only 197 such rows, so error checking wasn't too 
tedious. I went through each line to check if the two names
are really the same and deleted the lines where this was 
not the case. I can now confidently replace the names in one 
dataset from those in another.

```{r n_resolve}
resolved <- here::here("data/16th",
                       "conflicting_names_resolved.csv") %>%
            readr::read_csv(.) %>% arrange(mp_name.y)


```
Let us replace the names in the attributes data frame.

```{r}
replace_these <- fixed(resolved$mp_name.y)
replace_with  <- fixed(resolved$mp_name.x)

mps_attributes %<>% arrange(mp_name)
mps_attributes[mps_attributes$mp_name %in% replace_these, ] %<>%
mutate_at("mp_name", str_replace, pattern = replace_these,
                                  replacement = replace_with)

```

Now I can join by name, check the output and write it out.

```{r}
attendance_with_attributes <- left_join(all_sessions,
                                        mps_attributes,
                                        by = "mp_name")

dim(attendance_with_attributes)
dim(all_sessions)

```
That's what we expect to see with a left-join, so things are
looking good.

```{r}

here::here("data/16th", "attendance_with_attributes.csv") %>%
readr::write_csv(x = attendance_with_attributes, path = .)


here::here("data/16th", "sessions_durations.csv") %>%
  readr::write_csv(x = sessions_durations, path = .)


here::here("data/16th", "all_sessions_attendance.csv") %>%
  readr::write_csv(x = all_sessions, path = .)

```


